Synthesizing Quotient Lenses

### 2018-03-15 papers/QuotientLenses/main.tex
comment in the related work section
Polish abstract

\begin{document}
\title{Synthesizing Quotient Lenses}
\begin{abstract}
{\em Quotient lenses} are bidirectional transformations whose correctness
laws are ``loosened'' by specified equivalence relations, allowing
inessential details in concrete data formats to be suppressed. 
For example, a programmer could use a quotient lens to define 
a transformation that ignores the order of fields in XML data, so
that two XML files with the same fields but in different orders would be
considered the same, allowing a single, simple lens to handle them both.

Building on recent work that showed how to synthesize bijective lenses from high
level specifications, we offer three innovations to synthesize quotient lenses.
First, we introduce {\em quotient regular expressions} (QREs), that easily mark
aspects of data formats that are inessential and hence give a high-level
specification of regular languages with an equivalence relation defined on
them. Next, we introduce {\em QRE lenses}, i.e., lenses mapping between QREs.
Our key technical result is a proof that every QRE lens can be transformed into
one of the form $c ; \ell ; d^{-1}$, where $c$ and $d^{-1}$ are {\em
canonizers}, $\ell$ is a bijective transformation, and $;$ is
composition.  Intuitively, the canonizers normalize data by
picking a canonical representative of the relevant equivalence relations. This
theorem means we can push all normalization to the outer edges of a
bidirectional transformation, leaving in the middle a pure bijective
transformation. Finally, we leverage this normalization theorem to {\em
synthesize} QRE lenses from a pair of QREs and example input-output pairs,
reusing earlier work on synthesizing plain bijective lenses. We have
implemented QREs and QRE lens synthesis in the bidirectional programming
language Boomerang. We evaluate the effectiveness of our approach by using it
to synthesize QRE lenses between various real-world data formats in the
Optician benchmark suite.

@@

The \kw{Formless} type is given by
\begin{center}
\begin{lstlisting}
type Formlens a = Maybe a $\to$ [Int] $\to$ (Html,Env $\to$ Maybe a,[Int])
\end{lstlisting}
\end{center}
with the differences between this definition and the standard definition for
formlets being an extra \codefont{Maybe\; a} parameter that provides an
optional initial value of the form, the namesource being changed to a list of
integers, and the {\em collect} function being allowed to return an optional
value.

>>>>

\begin{document}
\title{Synthesizing Quotient Lenses}
\begin{abstract}
{\em Quotient lenses} are bidirectional transformations whose correctness
laws are ``loosened'' by specified equivalence relations, allowing
inessential details in concrete data formats to be suppressed. 
For example, a programmer could use a quotient lens to define 
a transformation that ignores the order of fields in XML data, so
that two XML files with the same fields but in different orders would be
considered the same, allowing a single, simple program to handle them both. 
%
Building on a recently published algorithm for synthesizing plain bijective
lenses from high-level specifications, we show how to synthesize bijective
{\em quotient} lenses in three steps.  First, we introduce {\em quotient regular
  expressions} (QREs), annotated regular expressions that conveniently mark
inessential aspects of string data formatsl; each QRE specifies,
simulteneously, a regular language and an equivalence relation on it.
Second, we introduce {\em QRE lenses}, i.e., lenses mapping between QREs.
Our key technical result is a proof that every QRE lens can be transformed into
one of the form $c ; \ell ; d^{-1}$, where $c$ and $d$ are {\em
canonizers}, $\ell$ is a bijective transformation, and $;$ is
composition\bcp{This seems not quite accurate.  It makes it look like
  canonizers are bidirectional things, but they aren't---at least, we
don't talk about them in that way (I suppose we could define the ``other
direction'' to be the identity, but we do not).  What we really mean are
left quotient and right quotient (and then we don't need the $^{-1}$).
Similarly, the meaning of 
``;'' here is not sequential composition of either lenses or QREs, which are
the only two meanings we give to this symbol in the paper.}.  
% Intuitively,
% these canonizers normalize data by 
% picking a canonical representative from the relevant equivalence
% relations. 
% BCP: Don't think this adds much, and it gets between the theorem and its use.
This theorem means we can push all normalization to the outer edges of a
composite quotient lens, leaving in the middle a pure bijective
transformation. Third, we leverage this normalization theorem to {\em
synthesize} QRE lenses from a pair of QREs and example input-output pairs,
reusing earlier work on synthesizing plain bijective lenses. We have
implemented QREs and QRE lens synthesis as an extension to the bidirectional programming
language Boomerang. We evaluate the effectiveness of our approach by 
synthesizing QRE lenses between various real-world data formats in the
Optician benchmark suite.

@@

The \kw{Formless}\saz{Is this ``Formless'' ``Formlens'' or ``Formlet''? The code
  below says ``Formlens'' but the earlier parts of the paragraph say ``Formlet''}type is given by
\begin{center}
\begin{lstlisting}
type Formlens a = Maybe a $\to$ [Int] $\to$ (Html,Env $\to$ Maybe a,[Int])
\end{lstlisting}
\end{center}
with the differences between this definition and the standard definition for
formlets being an extra \codefont{Maybe\; a} parameter that provides an
optional initial value of the form, the namesource being changed to a list of
integers, and the {\em collect} function being allowed to return an optional
value.
